---
title: "R Notebook"
output: html_notebook
---

Testing small sample dataset for replicability of perf methods 

```{r packages}

library(tidyverse)
library(readxl)
library(stringr)
library(lubridate)
library(scales)

```

```{r import}

dem <- read_xlsx("G:/My Drive/@DataSF Root/Consulting/MBO - PERF/911_Police_Data_FY2016-17_Final.xlsx", 
                 sheet = "DEM", 
                 skip = 2,
                 col_types = c("text",
                               "text",
                               "text",
                               "text",
                               "text",
                               "date",
                               "date",
                               "date",
                               "date",
                               "date",
                               "date",
                               "text",
                               "text"))

onview <- read_xlsx("G:/My Drive/@DataSF Root/Consulting/MBO - PERF/911_Police_Data_FY2016-17_Final.xlsx", 
                 sheet = "ONVIEW", 
                 skip = 2,
                 col_types = c("text",
                               "text",
                               "text",
                               "text",
                               "text",
                               "date",
                               "date",
                               "date",
                               "date",
                               "date",
                               "date",
                               "text",
                               "text"))

data <- bind_rows(dem,onview)
rm(dem,onview)


```

### Clean

- fix date/time 
- split and relabel dispatch 1st unit

Odd things.  The two_officer flag occasionally has a other characters besides 1 


```{r clean}


# fix date/time
data <- 
data %>% 

  # create date/time columns for entry, dispatch, and close
  #convert to text
  mutate(`ENTRY DATE` = as.character(`ENTRY DATE`),
         `ENTRY TIME` = as.character(`ENTRY TIME`),
         `DISPATCH DATE` = as.character(`DISPATCH DATE`),
         `DISPATCH TIME` = as.character(`DISPATCH TIME`),
         `CLOSE DATE` = as.character(`CLOSE DATE`),
         `CLOSE TIME` = as.character(`CLOSE TIME`)) %>% 
  #remove erroneous date info from time fields
  mutate(`ENTRY TIME` = str_replace(`ENTRY TIME`, "1899-12-31 ", ""),
         `DISPATCH TIME` = str_replace(`DISPATCH TIME`, "1899-12-31 ", ""),
         `CLOSE TIME` = str_replace(`CLOSE TIME`, "1899-12-31 ", "")) %>% 
  # combine date & time
  mutate(call_entry = str_c(`ENTRY DATE`,`ENTRY TIME`, sep = " "),
         call_dispatch = str_c(`DISPATCH DATE`, `DISPATCH TIME`, sep = " "),
         call_close = str_c(`CLOSE DATE`, `CLOSE TIME`, sep = " ")) %>% 
  #convert back to POSIXct
  mutate(call_entry = ymd_hms(`call_entry`),
         call_dispatch = ymd_hms(`call_dispatch`),
         call_close = ymd_hms(`call_close`)) 

#fix dispatch 1st unit
data <- 
data %>%
  mutate(
    #extract 1st two characters
    responding_unit = str_sub(`DISPATCH 1ST UNIT`, 1, 2),
    
    #extract 2nd to last character
    sector = str_sub(`DISPATCH 1ST UNIT`, 
                     (str_length(`DISPATCH 1ST UNIT`) -1), 
                     (str_length(`DISPATCH 1ST UNIT`) -1)),
    
    #extract last character
    shift = str_sub(`DISPATCH 1ST UNIT`, 
                    str_length(`DISPATCH 1ST UNIT`), 
                    str_length(`DISPATCH 1ST UNIT`)),
    
    #if dispatch 1st unit has 5 characters extract 3rd character, if not leave blank
    two_officers = if_else(str_length(`DISPATCH 1ST UNIT`)==5,
                           str_sub(`DISPATCH 1ST UNIT`, 3,3),
                           NA_character_)) 

# data check
data %>% distinct(DGROUP)
data %>% distinct(responding_unit)
data %>% distinct(sector)
data %>% distinct(shift)
data %>% distinct(two_officers)

data %>%
  ggplot(aes(responding_unit)) +
  geom_bar() +
  coord_flip()

data %>%
  distinct(two_officers)

data %>%
  #filter(two_officers != "1") %>% 
  ggplot(aes(two_officers)) +
  geom_bar() +
  coord_flip()

data %>%
  filter(two_officers != "1") %>% 
  ggplot(aes(two_officers)) +
  geom_bar() +
  coord_flip()

```
### Create variables
- same_district = did the call district match the dispatch unit district
- dur_sec = how many seconds between dispatch time close time
- dur_min = how many minutes between dispatch time close time
- dur_hrs = how many hrs between dispatch time close time
- duration = hr:min:sec


```{r variable creation}

data %>% 
  mutate(duration = call_close - call_dispatch) %>% 
  ggplot(aes(duration)) +
  geom_bar() +
  scale_y_continuous(labels = comma)

```

